---
title: "Yelp Review Analysis"
author: "Juliann McEachern"
date: "July 17, 2019"
output: 
  html_document:
    theme: paper
    highlight: pygments
    toc: true
    toc_float: true
    toc_depth: 2
    df_print: paged
    code_folding: show
---


# Final Project 

For the final project, we will be examining the relationship between yelp reviews and businesses in the metropolitan area of Phoenix, Arizona. The data was obtained from Kaggle's 2013 Yelp Review Challenge and was subsetted to include only businesses within the food and beverage industries.

We have a two-part goal in this assignment:

1.  Identify the links between key Yelp users and businesses within the Phoenix, Arizona community.  
2.  Evaluate the relationship between the text contents of reviews and the rating a business received. 

```{r dependencies, echo=F, warning=F, message=F, comment=F}
## data processing packages
library(tidyverse)

##formatting packages
library(knitr); library(kableExtra); library(default)

##network packages
library(igraph)

## knit sizing
options(max.print="100"); opts_knit$set(width=75) 


## augment chunk output
opts_chunk$set(echo=T,cache=F, tidy=T,comment=F,message=T,warning=T) #change message/warning to F upon completion


## set table style for consistency
default(kable) <- list(format="html")
default(kable_styling)  <- list(bootstrap_options = "hover",full_width=T, font_size=10)
default(scroll_box) <- list(width = "100%")
```

## Data Aquisition & Tidying {.tabset .tabset-fade .tabset-pills}

Data was acquired from Kaggle as a JSON file for a project conducted in [Data 612](http://rpubs.com/jemceach/D212-Final-Project). Our network uses the subsetted data from that project. This subset is stored as a csv file in our data folder and was read into this report for further review.  

We added additional transformations to meet our project goals and separated the data into seperate dataframes for network building and text processing. 

```{r aquisition-tidying, echo=F}
# load data
yelp<- read.csv("data/yelp.csv"); yelp <- select(yelp,-X, -business_id, -user_id) %>% rename(businessID = itemID)

# transform data for network analysis 
yelp_network <- yelp %>% 
  select(userID, businessID, name, stars, latitude, longitude) %>% 
  group_by(businessID) %>% 
  add_count() %>%
  rename(size = n)%>%
  ungroup() %>%
  group_by(userID) %>% 
  add_count() %>%
  rename(weight = n)%>%
  ungroup() 

# transform data for text processing 
yelp_reviews <- yelp %>% 
  transform(reviewID=match(review_id, unique(review_id))) %>% 
  select(-review_id, -longitude, -latitude) %>%
  select(userID, businessID, reviewID, name, city, state, votes.funny, votes.useful, votes.cool, 
         stars, text, date)
```

### Network Data 
```{r view-network-data, echo=F}
yelp_network
```


### Text Data 

```{r view-text-data, echo=F}
yelp_reviews
```


# Network Analysis

## 2-Mode Network  {.tabset .tabset-fade .tabset-pills}

We set up our initial 2-mode network by connecting our businesses and users using a weighted incidence matrix. We plotted our graph using the `plot.igraph` function and verified our network was created properly. 

### Build Network

```{r build-network}
# define edges; spread data from long to wide; convert to matrix
edges <- yelp_network %>% 
  select(businessID, userID, weight) %>% 
  spread(businessID, weight, fill = 0) %>% 
  column_to_rownames('userID') %>%
  as.matrix()

# define nodesets 
business_nodes <- yelp_network %>% 
  select(businessID, name, size) %>% 
  mutate(type = 'business', name = as.character(name)) %>% 
  distinct()

user_nodes <- yelp_network %>% 
  select(userID) %>% 
  mutate(name=paste0("U",userID),sizes = NA, type = 'user') %>% 
  distinct()

# bind rows
nodes <- bind_rows(business_nodes,user_nodes)

# initiate graph from matrix
g <- graph_from_incidence_matrix(edges, weighted=T)

# Define vertex color/shape
V(g)$shape <- ifelse(V(g)$type, "circle", "square")
V(g)$color <- ifelse(V(g)$type, "red", "white")
```

### Network Graph

```{r plot-graph, echo=F}
# Plot network
plot.igraph(g, 
            layout=layout.bipartite, 
            vertex.frame.color="black",
            vertex.label=NA)
```

### Verification

```{r verify-network, echo=F}
# verify vertices (F = User; T = Business)
node_count <-table(V(g)$type==T)

# Verify connections
rbind(Business.Nodes = toString(node_count[2]), 
      User.Nodes = toString(node_count[1]),
      Is.Weighted = toString(is.weighted(g)), 
      Is.Bipartite = toString(is.bipartite(g))) %>%
  kable(caption="Verify Node Counts and Connectivity") %>%
  kable_styling()
```

## Edge-Trimming  {.tabset .tabset-fade .tabset-pills}

To better understand our network, we applied the island method to see our most influential user and businesses within our dataset.



# Text Analysis

Possible reference: https://towardsdatascience.com/sentiment-analysis-with-text-mining-13dd2b33de27

# Conclusion

Final thoughts. 

----------

# References

Inspiration for this project was derrived from the following sources: 

1.  Data Source: https://www.kaggle.com/c/yelp-recsys-2013/data
2.  Related Project: http://rpubs.com/jemceach/D612-Final-Project
3.  R Network Reference: https://kateto.net/network-visualization


